---
title: "Therapy"
author: "Yang"
date: "April 2, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load data
```{r data}
# The data
Y<- cbind(c(28,4,41,12),c(45,12,44,7),c(29,5,20,3),c(26,2,20,1))
colnames(Y)<-c("ProgressiveDisease","NoChange","PartialRemission","CompleteRemission")
therapy <- c("S","S","A","A")
gender <- rep(c("M","F"),2)
```

#Baseline Categorical Logit Model

##Fit
```{r Baseline cat logit}
library(nnet)
#Progressive Disease as baseline
b1<-multinom(Y~as.factor(therapy))
beta1<-coefficients(b1)
summary(b1)
```

##Results
```{r}
result=predict(b1,type ="prob")
result=cbind(therapy,result)
#predicted probability
result
# proportional Therapy S vs A
S <-c(28,4, 45,12,29,5,26,2)
A <-c(41,12,44,7,20,3,20,1)
PS=sum(S)/(sum(S)+sum(A))
PA=sum(A)/(sum(S)+sum(A))
cbind(PS,PA)
```

##Interpretation
In our model with Progressive Disease as baseline:\newline
Under the alternating therapy treatment, we have $log(\frac{\pi_{N}}{\pi_{Prog}})= \beta_{11} = -0.03846754$; Under the sequential therapy treatment, we have $log(\frac{\pi_{N}}{\pi_{Prog}}) = \beta_{11} + \beta_{12} = -0.03846754 + 0.6157805 = 0.577313$.\newline 
The log odds ratio: $log\frac{odd_S}{odd_A}=log(\frac{\pi_{N}}{\pi_{Prog}}|S)-log(\frac{\pi_{N}}{\pi_{Prog }}|A)= \beta_{12}=0.6157805$, which suggests the odds of "Progressive Disease" vs "No Disease" with alternating sequential and alternating therapy are $\frac{odd_S}{odd_A}=e^{0.6157805}= 1.851101$, meaning sequential therapy are better than alternating therapy.

##Confidence Interval
```{r}
SE <- summary(b1)$standard.errors
#95% CI
CI.u <- exp(beta1[1,2]+ 1.96*SE[1,2])
CI.l <- exp(beta1[1,2]- 1.96*SE[1,2])
cbind(CI.l,CI.u)
```
From the confidence interval, we can also conclude that we are 95% confident that sequential therapy has a better treament effect than alternating treament effect since 1 is not trapped by our confidence interval.



#Cummulative Logit Model

##Fit
```{r fit Y~therapy}
library(VGAM)
c1=vglm(Y~therapy,propodds(reverse=F))
summary(c1)
```

##Coefficients and interpretation
Exponentiated coefficients:therapyS $e^\beta=e^{-0.5699}=0.5655741$. Interpretation: $log\frac{\frac{Pr(Y\leq j|Therapy=S)}{Pr(Y\geq j|Therapy=S)}}{\frac{Pr(Y\leq j|Therapy=A)}{Pr(Y\geq j|Therapy=A)}}=\beta \implies \frac{\frac{Pr(Y\leq j|Therapy=S)}{Pr(Y\geq j|Therapy=S)}}{\frac{Pr(Y\leq j|Therapy=A)}{Pr(Y\geq j|Therapy=A)}}=e^\beta=e^{-0.5699}=0.5655741$. Since 0.5655741 is smaller than 1, we can conclude that sequential therapy has a better effect than alternating therapy.

##Confidence Interval
```{r CI}
e=exp(1)
e^confint(c1)
```
95% confidence interval for exponentiated coefficients of therapy:[0.3735997,  0.8561948]. Interpretation: we are 95% confident that the true odds ratio for the effect of therapy fall within [0.3735997,  0.8561948].

#Should Gender Be Included in the Models Above?

##Baseline Categorical Model
```{r baseline}
library(nnet)
b2<-multinom(Y~therapy+gender)
beta2<-coefficients(b2)
summary(b2)
anova(b1,b2,test="Chi")
```
According to the anova table, the difference between baseline category logit model with therapy and the one with therapy+gender is not statistically significant. Therefore, we would not include gender as a predictor.

##Cummulative Logit Model
```{r cummulative logit model}
c2<-vglm(Y~therapy+gender,propodds(reverse=F))
summary(c2)
p_val <- pchisq(2*(logLik(c2)-logLik(c1)),df=df.residual(c1)-df.residual(c2), lower.tail=FALSE)
p_val
```
The p value from the test is 0.057<0.05, from which we can conclude that gender should not be included in the cummulative logit model.

##Conclusion
Therefore, we conclude that gender should not be included in the model(s).